---
title: "Day7"
format: html
editor: visual
execute: 
  warning: false
---

```{r}
library(tidyverse)
library(palmerpenguins)
```

## Data wrangling refresher

1.  starting from penguins, write a single pipe sequence in which you:
    -   **Only** include penguins at Dream and Biscoe lslands
    -   Remove the year and sex variables
    -   Add a new column body_mass_kg with penguin mass converted from g to kg
    -   Rename the island variable to location

```{r}
penguins %>% filter (island  %in% c("Dream","Biscoe")) %>%
  select (-year,-sex) %>%
  mutate(body_mass_kg = body_mass_g * 0.001) %>%
  rename(location = island)
```

2.  Starting from penguins, write a single piped sequence to:

    -   Limit to Adelie penguins
    -   Remove any observations where filpper_length_mm is NA(!is.na()) or drop_na)
    -   Group the data by sex
    -   Create a summary table that contains the mean, standard deviation and sample size of flipper length for male and female Adelie penguins.

    ```{r}
    adelie_summary <- penguins %>% 
      filter(species == "Adelie") %>%
      drop_na(flipper_length_mm)%>%
      group_by(sex) %>%
      summarize(mean_flipper_length = mean(flipper_length_mm), sd_flipper_length = sd(flipper_length_mm),sample_size =n())

    # filter(speices == "Adelie", !is.na(filler_length_mm)) <- include drop_na() when you want to remove one of column
    ```

```{r}
penguins %>%
  group_by(species, island, sex) %>%
  summarize(sample_size = n())

# Count() is just wrapping together group_by() + summarize() + n() to get counts of obsevations. 
penguins %>% count (species, island, sex)
```

```{r}
animals <- data.frame(
  stringsAsFactors = FALSE,
          location = c("lagoon", "bluff", "creek", "oaks", "bluff"),
           species = c("bobcat", "coyote", "fox", "squirrel", "bobcat"),
          maturity = c("adult", "juvenile", "adult", "juvenile", "adult")
)
```

```{r}
sites <- data.frame(
  stringsAsFactors = FALSE,
          location = c("beach", "lagoon", "bluff", "oaks"),
    full_site_name = c("Goleta Beach","UCSB Lagoon",
                       "Ellwood Mesa","Fremont Campground"),
      jurisdiction = c("SB City", "UCSB", "SB City", "USFS")
)
```

## full_join(x,y):

```{r}
full_join_example <- full_join (animals, sites)
```

## left_join(x,y):

```{r}
left_join_example <- left_join (animals, sites)
```

## inner_join(x,y): 

```{r}
inner_join_example <- inner_join (animals, sites)
```
